---
title: "Line Diagrams"
linktitle: "Line Diagrams"
output:
  blogdown::html_page:
    toc: true
    toc_depth: 1
menu:
  example:
    parent: Examples
    weight: 1
type: docs
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why would I need code for a line diagram?

Line diagrams are easy to draw by hand, but can get unwieldy for real datasets. But sometimes plotting the line diagram can be useful. This includes settings when you want a quick visual check of

- the administrative censoring date
- the presence of late entry
- the maximum follow-up time
- etc


## R code to produce line diagrams

### Simple simulated data (used in L1)

First, we generate simple data on 20 participants as describe by L1:

> "Say you wish to estimate the 5-year risk of death among people entering HIV care. You have a database of people entering HIV care between 2012 and 2020."


```{r message=FALSE, warning=FALSE}
### Generate some data -----
require(dplyr)
set.seed(123)
year0 <- runif(20, min=2012, max = 2020)
t <- runif(20, min=2, max = 15)
dat <- data.frame(year0, t)


dat <- dat %>% mutate(y=ifelse(t+year0>2020 | t>5, 0, 1),
                      t = ifelse(t+year0>2020, 2020-year0, t), 
                      t = ifelse(t>5, 5, t),
                      id = row_number())

dat
```

Produce the line diagram by creating line segments and points in `ggplot`.

```{r message=FALSE, warning=FALSE}

# Calendar time as timescale
library(grid)
library(ggthemr)
ggthemr('solarized')
line <- ggplot() +
  geom_segment(data = dat %>% filter(y==0), aes(x = year0, y = id, xend = year0+t, yend = id),  arrow = arrow(length = unit(0.1, "cm"))) +
  geom_segment(data = dat %>% filter(y==1), aes(x = year0, y = id, xend = year0+t, yend = id)) +
  scale_y_continuous(name = "ID", breaks = c(1, 5, 10, 15, 20), limits = c(0,20))+
  scale_x_continuous(name = "Calendar Time", breaks=c(2012, 2014, 2016, 2018,  2020), limits = c(2012, 2021)) +
  geom_point(data = dat %>% filter(y==1), aes(x = year0+t, y = id), color = "red", size = 0.6) +
  theme(text = element_text(size = 14, family = "Open Sans"))

line
```

Reorganize plot to show time since entry into HIV care on the `x`-axis.

```{r message=FALSE, warning=FALSE}

# Calendar time as timescale
line2 <- ggplot() +
  geom_segment(data = dat %>% filter(y==0), aes(x = 0, y = id, xend = t, yend = id),  arrow = arrow(length = unit(0.1, "cm"))) +
  geom_segment(data = dat %>% filter(y==1), aes(x = 0, y = id, xend = t, yend = id)) +
  scale_y_continuous(name = "ID", breaks = c(1, 5, 10, 15, 20), limits = c(0,20))+
  scale_x_continuous(name = "Calendar Time", breaks=c(0, 1, 2, 3, 4, 5), limits = c(0, 6)) +
  geom_point(data = dat %>% filter(y==1), aes(x = t, y = id), color = "red", size = 0.6) +
  theme(text = element_text(size = 14, family = "Open Sans"))

line2
```

### Example data (similar to Cole & Hudgens 2010)

Next, we read in some sample data similar to that used in Cole & Hudgens 2010.

```{r echo=FALSE, message=FALSE, warning=FALSE}
### Read in data -----
library(grid)
library(ggthemr)
ggthemr('solarized')
exdat <- read.csv("../../static/data/exdat.csv", header = TRUE)
head(exdat)

```

### Calendar timescale

```{r, warning=FALSE}

### Create second dataset with only events ------
exdat2 <- exdat[exdat$d == 1, ]

### Plot the lines -----
line <- ggplot(data = exdat) +
  geom_segment(aes(x = year, y = newid, xend = yearw, yend = newid), lty = "dotted") +
  geom_segment(aes(x = yearw, y = newid, xend = yeart, yend = newid)) +
  ylab("ID") +
  scale_x_continuous(name = "Calendar Time", breaks=c(2010, 2015,  2020)) +
  geom_point(data = exdat2, aes(x = yeart, y = newid), color = "red", size = 0.5) +
  theme(text = element_text(size = 16, family = "Open Sans"))

line

```


### Time since AIDS diagnosis

```{r, warning=FALSE}
line2 <- ggplot(data = exdat) +
  geom_segment(aes(x = 0, y = newid, xend = w, yend = newid), lty = "dotted") +
  geom_segment(aes(x = w, y = newid, xend = t, yend = newid)) +
  ylab("ID") +
  scale_x_continuous(name = "Time since AIDS diagnosis", breaks = c(0,2, 4, 6, 8, 10)) +
  geom_point(data = exdat2, aes(x = t, y = newid), color = "red", size = 0.5) +
  
  theme(text = element_text(size = 14, family = "Open Sans"))

line2
```

